---
title: "Exam 2"
author: "Daniel Mahoney"
date: "6/26/2020"
output: word_document
---

```{r}
#clear the environment
rm(list=ls(all=TRUE))
#load library of rio
library(rio)
inequality_data <- import("/Users/Owner/Downloads/inequality.xlsx")
```

This data set is cross-sectional. A cross-sectional dataset, is data taken at a specific point in time. Our data-set works with various countries during the year 2015, and inequality numbers that can be taken at that specific point in time. We can tell by looking at some of the data.

```{r}
#check dataset
head(inequality_data)
```

```{r}
#check subsets of inequality_gini scores Denmark and Sweden
subset(inequality_data, country == "Denmark" | country == "Sweden", select= c(country, inequality_gini))
```

```{r}
#check subsets of inequality_gini scores of Brazil
subset(inequality_data, country == "Brazil", select= c(country, inequality_gini))
```

It is better to have a low inequality_gini score, as the countries with lesser inequality, have lower inequality_gini scores.

```{r}
#check dataset
head(inequality_data)
```

```{r}
#remove accents
remove.accents <- function(s) {
  # character substitution for Belarus
  old1 <- "ú"
  new1 <- "u"
  s1<- chartr(old1,new1,s) 
}
inequality_data$country <- remove.accents(inequality_data$country)
head(inequality_data)
```

```{r}
#sort the data frame
inequality_data <- inequality_data[order(inequality_data$inequality_gini),]
head(inequality_data)
```

```{r}
#stargazer library
library(stargazer)
#find mean of inequality_gini
stargazer(inequality_data, type = "text")
```

The mean of the inequality_gini scores is 36.814.

```{r}
#setting high_inequality and low_inequality
high_inequality<- c(inequality_data$inequality_gini) > 36.814
low_inequality<- c(inequality_data$inequality_gini) < 36.814
#if else test
ifelse(test = high_inequality ==TRUE, yes = "0", no ="1")
```

```{r}
library(gmodels)
#cross tab high_inequality and low_inequality
crosstab(inequality_data$inequality_gini ,row.vars = "Country", col.vars = "number of observations", type="mean")
```

```{r}
#for loop to print names
names <- c("The World Bank", "African Development Bank", "Bill and Melinda Gates Foundation")
for(x in names) {
print(x)
}
```

A variable that I find to be correlated with inequality, is the variable of "Mortality rate, under-5(per 1,000 live births). This variable seems to show that those with a high mortality rate under 5, also tend to have higher inequality_gini numbers. This would seem to suggest that countries with worse medical care and insitutuions, would also have a higher quantity of inequality.

```{r}
#import data
remotes::install_github('vincentarelbundock/WDI')
mortality_rate<-population_data =WDI(country = "all",indicator =c("SH.DYN.MORT"),start = 2015, end = 2015, extra = FALSE, cache = NULL)

```

```{r}
library(tidyverse)
#merge datasets
merged_df <- left_join(inequality_data, mortality_rate, by=c("country", "year") )
```



```{r}
#get rid of NA in inequality_data
library(tidyverse)
merged_df <- 
  inequality_data %>% 
  dplyr::filter(!(inequality_data$inequality_gini=="NA"))
```

```{r}
#create data_greater_30
library(tidyverse)
data_greater_30<- 
  inequality_data %>% 
  dplyr::filter(inequality_data$inequality_gini > 30)
```

```{r}
#find sum in inequality_gini of data_greater_30
my.data<- matrix(data_greater_30$inequality_gini)
apply(my.data, 2, sum)
```
```{r}
# label the variables of merged_df
#install.packages("labelled")
library(labelled)
var_label(merged_df) <-list(`country`= "Country",`year`= "year",`inequality_gini`= "mortality rate under 5")
# save the dataset in Stata format with the labels
library(rio)
export(merged_df, file = "final_data.dta")
```